---
title: "Insect Abundance Seasonal ARIMA"
author: "Isabella Escamilla"
output:
    html_document:
      toc: true
      toc_float: true
      code_folding: show
---

```{r, include=FALSE}
library(dplyr)
library(astsa)
library(MASS)
library(reshape2)
library(reshape)
library(tidyverse)
```

## Introduction
The goal of this analysis is to create a seasonal ARIMA model of the insect abundance data set. First, I will import the data and make any modifications necessary.

```{r}
dataloc <- './DataSource/data/InsectAbundanceBiomassData.csv'
data <- read.csv(dataloc)
summary(data)
```

There's a lot of data from different sources, so we will choose one data source to examine for now to avoid inconsistencies. We will be look at the grasslands butterflies in Wisconsin.

```{r}
data2 <- data %>% filter(DataSource_ID == 1476)
head(data2)
```

The months are inconsistent and won't work as a time series object, so we will aggregate by year.

```{r}
data2 <- na.omit(data2)
agg_tbl <- data2 %>% group_by(Year) %>% 
  summarise(mean_num=mean(Number),
            .groups = 'drop')
agg_tbl
```

Now we can turn it into a time series.

```{r}
butterfly_ts <- ts(agg_tbl$mean_num, start = 1987, end = 2017)
plot.ts(butterfly_ts)
```

We can see that the time series needs to be detrended.

```{r}
diff <- diff(butterfly_ts)
plot.ts(diff)
```

Now we'll look at the ACF.

```{r}
acf2(diff)
```

Both the ACF and PACF cut off after lag 0, indicating that this likely follows a white noise model. This means that p and q are equal to zero, and d is equal to 1 since we took the first difference.

```{r}
sarima(butterfly_ts, 0, 1, 0, 0, 0, 0)
```

The model's estimate is significant, so we will try making predictions for the next 6 years.

```{r}
sarima.for(butterfly_ts, 6, 0, 1, 0, 0, 0, 0)
```

### Questions
One of the reasons I chose this data set was because it included months/seasons, but I'm not sure what to do when it doesn't include every month and it's inconsistent on which months it includes. I would like to use that information so I could fit an actual SARIMA model, but I'm not sure how.